Scaffolder - Software for \change{Reproducible}{Manual} Genome Scaffolding.

### 2011-08-15 src/article.tex
Revise Background

The advent of high\add{-}throughput sequencing methods has led to a renewed
interest in algorithms to solve the problem of genome assembly
\cite{pop2008,pop2009}. \remove{Nonetheless, due to }The complexity of merging
large numbers of overlapping reads \add{can lead to} genome assembly software
may be unable to produce a complete sequence. Instead, the algorithm may
generate several large assembled regions of sequence (`contigs') composed of
the many individual reads\remove{ that represent assembly into the longest
possible sequences, given the data. Nonetheless, }These contigs represent
a fragmented picture of the genome and \add{therefore} require additional work
to join together into a complete sequence. \pb

@@

The process of \change{resolving a genome sequence and building a}{finishing
a genome} scaffold \change{may} use wet laboratory methods, \emph{in silico}
methods, or a combination of both. An example of a computational method might
use the \remove{existing} paired-read data from the sequencing stage.
\remove{where reads were produced in pairs at a known distance apart in the
genome.} The occurrence of paired reads in separate contigs can be used to
\change{probabilistically estimate}{estimate probabilistically} the order and
distance between these contigs. Alternatively, laboratory methods may use PCR
to amplify the unknown DNA in a gap region, then use \remove{traditional}
Sanger sequencing to determine the sequence of this gap. Computational methods
are \remove{more} preferable as they are less costly in laboratory time and
materials compared to manual gap resolution, and use the available data that
exists following sequencing \cite{nagarajan2010}. Finally when the scaffold
cannot be completely resolved, \emph{in silico} software packages exist to
suggest the likely primers necessary for PCR amplifying the sequence in gap
regions \cite{gordon2001}. \pb

Examples of \emph{in silico} methods include \remove{using synteny to}
comparing the assembled contigs to a complete reference genome sequence
\change{by}{to} search\remove{ing} for areas of sequence similarity between the
two. Any areas of corresponding sequence with the reference genome can be used
to infer contig placement and build the contigs into a scaffold
\cite{richter2007,zhao2008,assefa2009}. Recombination within each of the two
compared genomes can, however, reduce the efficacy of this. \pb

Repeat regions may also be responsible for multiple gaps when building
a genome sequence; tandemly repeated nucleotide regions in the genome produce
multiple reads with similar sequence. As many assembly algorithms rely on
sequence overlaps between reads to build a contig, the similarity between
repeat-region reads can result in the assembly collapsing into an artificially
short sequence or being ignored by more conservative assembly algorithms. Such
regions can be resolved by using algorithms \add{which} specifically reassemble
the collapsed repeat region correctly \cite{mulyukov2002,koren2010}.
\change{while a}{A} related approach uses unassembled sequence reads matching
the regions around a scaffold gap to construct a uniquely overlapping set of
reads across \change{it}{the gap}. \cite{tsai2010}. \pb

>>>>

The advent of high\add{-}throughput sequencing methods has led to a renewed
interest in algorithms to solve the problem of genome assembly
\cite{pop2008,pop2009}. \remove{Nonetheless, due to }The complexity of merging
large numbers of overlapping reads \add{can lead to} genome assembly software
\remove{may} being unable to produce a complete sequence. Instead, the
algorithm may generate several large assembled regions of sequence (`contigs')
composed of the many individual reads\remove{ that represent assembly into the
longest possible sequences, given the data. Nonetheless, }These contigs
represent a fragmented picture of the genome and \add{therefore} require
additional work to join together into a complete sequence. \pb

@@

The process of \change{resolving a genome sequence and building a}{finishing
a genome} scaffold \remove{may} uses wet laboratory methods, \emph{in silico}
methods, or a combination of both. An example of a computational method might
use the \remove{existing} paired-read data from the sequencing stage.
\remove{where reads were produced in pairs at a known distance apart in the
genome.} The occurrence of paired reads in separate contigs can be used to
\change{probabilistically estimate}{estimate probabilistically} the order and
distance between these contigs. Alternatively, laboratory methods may use PCR
to amplify the unknown DNA in a gap region then use \remove{traditional}
Sanger sequencing to determine the sequence of this gap. Computational methods
are \remove{more} preferable as they are less costly in laboratory time and
materials compared to manual gap resolution, and use the available data that
exists following sequencing \cite{nagarajan2010}. Finally when the scaffold
cannot be completely resolved, \emph{in silico} software packages exist to
suggest the likely primers necessary for PCR amplifying the sequence in gap
regions \cite{gordon2001}. \pb

Examples of \emph{in silico} methods include \remove{using synteny to}
comparing the assembled contigs to a complete reference genome sequence
\change{by}{to} search\remove{ing} for areas of sequence similarity between the
two. Any areas of corresponding sequence with the reference genome can be used
to infer contig placement and build the contigs into a scaffold
\cite{richter2007,zhao2008,assefa2009}. \add{Genomic} recombination
\remove{within each of the two compared genomes} can however reduce the
\change{efficiency}{efficacy} of this. \pb

Repeat regions may also be responsible for multiple gaps when building
a genome sequence; tandemly repeated nucleotide regions in the genome produce
multiple reads with similar sequence. As many assembly algorithms rely on
sequence overlaps between reads to build a contig, the similarity between
repeat-region reads can result in the assembly collapsing into an artificially
short sequence or being ignored by more conservative assembly algorithms. Such
regions can be resolved by using algorithms \add{that} specifically reassemble
the collapsed repeat region \remove{correctly} \cite{mulyukov2002,koren2010}.
\change{while a}{A} related approach uses unassembled sequence reads matching
the regions around a scaffold gap to construct a uniquely overlapping set of
reads across \change{it}{the gap}. \cite{tsai2010}. \pb
